---
title: "DavFinalProj"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#libraries
library(tidyverse)
library(rgdal)        
library(ggplot2)
library(wesanderson)
library(RColorBrewer)
library(ggmap)

```

Problem Understanding: 
Until now gender statistics have been region neutral- they show LFPR, wage gaps etc country-wise or region-wise, but without any exploration into causal relationships behind trends. If there is a huge difference in the manner in which fertility rates & labor force participation rates are related in sub-saharan africa as opposed to norwegian countries, we don't pause to wonder why that could be. Throwing light into causal relationships if any, into how women belonging to different regions react to maternity policies can give us insights on how to improve labor force participation rates with strategic interventions depending on the regions. 

Graphic1: that shows how every indicator across the world has improved for women, except for LFPR

Graphic 2: (aim is to stress need to address this issue urgently. So, Before Graphic3, show how world economic growth has been slowing down and overlap this with a fall in LFPR as well)

Graphic3: that shows current GDP @ current LFPR vs GDP with LFPR at full potential

Graphic 4: Plot Rate of fall of LFPR in different zones: North America, Sub-Saharan Africa, South Asia, Western Europe, East Asia

Research Q: it is evident from Graphic3 that increasing women LFPR is vital to global economic interests. Why then, despite increasing standard of living, healthcare, opportunities and laws for ensuring safety, does LFPR fall? How different are the scenarios between developed and developing countries? Using bayesian modeling, we try to explore causalities between FLFPR and other development indicators: maternity leave, per capita income, education levels, fertility rates, percentage of wage compensated during maternity leave etc. 
Using our inference, we provide policy recommendations for possible interventions. 

###########################################################################        
Graphic1 : 
major indicators: female literacy, maternal mortality, female life-expectancy, Net primary enrollment rate (%), Net secondary enrollment rate (%), 

STEP 2: data set exploration
```{r}
#import all relevant data dets and see heads.
setwd("~/Documents/GitHub/dav_final/GenderStatsData")
high_income=read.csv("flfpr_highincome.csv",sep=',',header=TRUE)
low_income=read.csv("flfpr_lowincome.csv",sep=',',header=TRUE)
pop= read.csv("population_total.csv", sep=',',header=TRUE, skip = 4)
ml_length <- read.csv("ml_length.csv",sep=',',header=TRUE,skip=2)
seatprop <- read.csv("seatproportion_national.csv",sep=',',header = TRUE,skip=2)
ml_wagespaid <- read.csv("ml_wagespaid.csv",sep=',',header=TRUE,skip = 2)
genderquota <- read.csv("voluntary_genderquota_parl.csv",sep=',',header=TRUE,skip=2)
#Share of female science technology engineering and mathematics graduates at tertiary level	
ger_tertiary <- read.csv("ger_tertiary.csv",sep=',',header=TRUE,skip=2)  
ger_secondary <- read.csv("ger_secondary_edu.csv",sep=',',header=TRUE)  
flfpr <- read.csv("FLFPR.csv",sep=',',header=TRUE,skip=4)
gdp <- read.csv("gdp.csv",sep=',',header=TRUE,skip=4)
gni <- read.csv("gni.csv",sep=',',header=TRUE,skip=4)

#selecting years 2000 to 2017 
#for [seatprop, ger_tertiary, ger_secondary], [flfpr, gdp, gni]
inds=c("seatprop", "ger_tertiary", "ger_secondary")
drops <- c("Age","Location","Occupation","LowerBound","UpperBound","Unit","Indicator.Name","OriginData","NatureData","Sex","Region","Country.Code")
years= c(2000:2017)
# for (i in 1:length(inds)) {
#     df=get(inds[i])
#     df=df[ , !(names(df) %in% drops)][,1:5]
# }
ger_tertiary=ger_tertiary[ ,!(names(ger_tertiary) %in% drops)][,1:3]
ger_secondary=ger_secondary[ , !(names(ger_secondary) %in% drops)][,1:3]
seatprop=seatprop[ , !(names(seatprop) %in% drops)][,1:3]
#doing this inside the forloop somehow doesn't work. will try again later.
ger_tertiary=ger_tertiary[which(ger_tertiary$Year>=2000),]
ger_tertiary=ger_tertiary[which(ger_tertiary$Year<2018),]
seatprop=seatprop[which(seatprop$Year>=2000),]
seatprop=seatprop[which(seatprop$Year< 2018),]
ger_secondary=ger_secondary[which(ger_secondary$Year>=2000),]
ger_secondary=ger_secondary[which(ger_secondary$Year<2018),]
colnames(ger_secondary)=c("Country","Year","indicator")
colnames(ger_tertiary)=c("Country","Year","indicator")
colnames(seatprop)=c("Country","Year","indicator")


#selecting years for [flfpr, gdp, gni]
drop_cols=c("Indicator.Name","Indicator.Code","Country.Code")
#why doesn't the for-loop work? doing it manually. -.-
flfpr=flfpr[,!names(flfpr) %in% drop_cols][,-c(4:42,61,62)]
flfpr=flfpr[,!names(flfpr) %in% c("X1960")]
gdp=gdp[,!names(gdp) %in% drop_cols][,-c(3:42,61,62)]
gni=gni[,!names(gni) %in% drop_cols][,-c(3:42,61,62)]
pop=pop[,!names(pop) %in% drop_cols][,-c(3:42,61,62)]

colnames(flfpr)=c("Country.Name","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017")
colnames(gdp)=c("Country.Name","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017")
colnames(gni)=c("Country.Name","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017")
colnames(pop)=c("Country.Name","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017")

flfpr=flfpr[1:(length(flfpr)-1)]
gdp=gdp[1:(length(gdp)-1)]
gni=gni[1:(length(gni)-1)]
pop=pop[1:(length(pop)-1)]
#have to melt the last three so that years are rows instead of separate columns. 
library(reshape2)

#can't be done in a forloop, because you have to store the molten dataframes under different names

molten.flfpr=melt(flfpr, id.vars = c("Country.Name"))
molten.gdp=melt(gdp, id.vars = c("Country.Name"))
molten.gni=melt(gni, id.vars = c("Country.Name"))
molten.pop=melt(pop, id.vars = c("Country.Name"))

colnames(molten.flfpr)=c("Country","Year","indicator")
colnames(molten.gdp)=c("Country","Year","indicator")
colnames(molten.gni)=c("Country","Year","indicator")
colnames(molten.pop)=c("Country","Year","indicator")

#the remaining: [genderquota, ml_wagespaid, ml_length]
ml_wagespaid1=ml_wagespaid[,!names(ml_wagespaid) %in% c("Indicator.Name","Footnote","Measure.Name","Data.Unit","Country.Code","Region","Measure.Value")]
colnames(ml_wagespaid1)=c("Country","indicator")

genderquota1=genderquota[,!names(genderquota) %in% c("Indicator.Name","Footnote","Measure.Name","Data.Unit","Country.Code","Region","Year")]
head(genderquota1)
colnames(genderquota1)=c("Country","indicator")

ml_length1=ml_length[,!names(ml_length) %in% c("Indicator.Name","Footnote","Measure.Name","Data.Unit","Country.Code","Measure.Value","Region")]
colnames(ml_length1)=c("Country","indicator")


#choosing countries we need from HighIncome and LowIncome countries
high_income_countries=subset(high_income,select=("Country.Name"))
low_income_countries=subset(low_income,select=("Country.Name"))
req_countries=rbind(high_income_countries,low_income_countries)
a=unique(req_countries$Country.Name)

#function to select countries
# SelectCountries <- function(df) {
#         # Remove entries from regions of aggregated countries
#         df <- subset(df,df$Country %in% a)
#         return(df)
# }
# indicators=c("seatprop","ger_tertiary","ger_secondary","molten.flfpr","molten.gdp","molten.gni","molten.pop","ml_length1","ml_wagespaid1","genderquota1")
# for(i in (1:length(indicators))){
#    SelectCountries(get(indicators[i]))
# }

seatprop=subset(seatprop,seatprop$Country %in% a)
indicator=rep("seatprop",length(seatprop[,1]))
seatprop=cbind(seatprop,indicator)

ger_tertiary=subset(ger_tertiary,ger_tertiary$Country %in% a)
indicator=rep("ger_tertiary",length(ger_tertiary[,1]))
ger_tertiary=cbind(ger_tertiary,indicator)

ger_secondary=subset(ger_secondary,ger_secondary$Country %in% a)
indicator=rep("ger_secondary",length(ger_secondary[,1]))
ger_secondary=cbind(ger_secondary,indicator)

molten.flfpr=subset(molten.flfpr,molten.flfpr$Country %in% a)
indicator=rep("molten.flfpr",length(molten.flfpr[,1]))
molten.flfpr=cbind(molten.flfpr,indicator)

molten.gdp=subset(molten.gdp,molten.gdp$Country %in% a)
indicator=rep("molten.gdp",length(molten.gdp[,1]))
molten.gdp=cbind(molten.gdp,indicator)

molten.gni=subset(molten.gni,molten.gni$Country %in% a)
indicator=rep("molten.gni",length(molten.gni[,1]))
molten.gni=cbind(molten.gni,indicator)

molten.pop=subset(molten.pop,molten.pop$Country %in% a)
indicator=rep("molten.pop",length(molten.pop[,1]))
molten.pop=cbind(molten.pop,indicator)

ml_length1=subset(ml_length1,ml_length1$Country %in% a)
indicator=rep("ml_length1",length(ml_length1[,1]))
ml_length1=cbind(ml_length1,indicator)

ml_wagespaid1=subset(ml_wagespaid1,ml_wagespaid1$Country %in% a)
indicator=rep("ml_wagespaid1",length(ml_wagespaid1[,1]))
ml_wagespaid1=cbind(ml_wagespaid1,indicator)

genderquota1=subset(genderquota1,genderquota1$Country %in% a)
indicator=rep("genderquota1",length(genderquota1[,1]))
genderquota1=cbind(genderquota1,indicator)

str_of_dataframes=c("seatprop","ger_tertiary","ger_secondary","molten.flfpr","molten.gdp","molten.gni","molten.pop")
main_df = data.frame()

    for (indicators in str_of_dataframes) {
        main_df = rbind(main_df, get(indicators))
    }

main_df=subset(main_df,main_df$Country %in% a)
colnames(main_df)=c("Country","Year","value","variable")
main_nona= main_df[!is.na(main_df$value),]
sum(is.na(main_nona$Value))

library(reshape2)
main_cast=dcast(main_nona, Country+Year~variable, sum)
main_cast_mean=dcast(main_nona, Country+Year~variable, mean)
write.csv(main_cast_mean,"main_cast.csv")

main_cast_ordered= main_cast_mean[order(main_cast_mean$Year),]

#mean by year 
mean_by_year=aggregate(main_cast_ordered[, 3:9], list(main_cast_ordered$Year), mean,na.rm=TRUE, na.action=NULL)
head(mean_by_year)

#normalising the dataset from [-1,1]
normFunc <- function(x) { (x-mean(x, na.rm = T))/sd(x, na.rm = T) }

mean_by_year[2:8] <- apply(mean_by_year[2:8], 2, normFunc)

#removing population
mean_by_year_final= mean_by_year[,-c(8)]
mean_by_year_final$Group.1=as.numeric(mean_by_year_final$Group.1) #the year column was still character
colnames(mean_by_year_final)=c("Year","seatprop", "ger_tertiary", "ger_secondary", "flfpr", "gdp", "gni")
head(mean_by_year_final)


#type+year~variable,sum
#number of countries are the same! 
#now we proceed to melt again

#correlation plot 
library(ggplot2)
library(Hmisc)

library(corrplot)
corr1 = rcorr(as.matrix(mean_by_year_final))
corrplot(corr1$r, is.corr = FALSE, method = "circle")

#-------------------------------------------------------------------------------------
#-------------------------------------------------------------------------------------

#now dividing the countries to high and low income and plotting correlations separately
main_cast_low = subset(main_cast_mean,main_cast$Country %in% low_income_countries$Country.Name)
main_cast_high = subset(main_cast_mean,main_cast$Country %in% high_income_countries$Country.Name)

#THERE SHOULD BE A WAY TO AUTOMATE THIS -.-
main_high_order= main_cast_high[order(main_cast_high$Year),]
main_low_order= main_cast_low[order(main_cast_low$Year),]

#mean by year 
mean_by_year_high=aggregate(main_high_order[, 3:9], list(main_high_order$Year), mean,na.rm=TRUE, na.action=NULL)
mean_by_year_low=aggregate(main_low_order[, 3:9], list(main_low_order$Year), mean,na.rm=TRUE, na.action=NULL)

#normalising the dataset from [-1,1]
normFunc <- function(x) { (x-mean(x, na.rm = T))/sd(x, na.rm = T) }

mean_by_year_high[2:8] <- apply(mean_by_year_high[2:8], 2, normFunc)
mean_by_year_low[2:8] <- apply(mean_by_year_low[2:8], 2, normFunc)


#removing population
mean_by_year_high= mean_by_year_high[,-c(8)]
mean_by_year_low= mean_by_year_low[,-c(8)]

mean_by_year_high$Group.1=as.numeric(mean_by_year_high$Group.1) #the year column was still character
mean_by_year_low$Group.1=as.numeric(mean_by_year_low$Group.1) #the year column was still character

colnames(mean_by_year_low)=c("Year","seatprop", "ger_tertiary", "ger_secondary", "flfpr", "gdp", "gni")
colnames(mean_by_year_high)=c("Year","seatprop", "ger_tertiary", "ger_secondary", "flfpr", "gdp", "gni")

#correlation plots for high and low income separately

corr_high = rcorr(as.matrix(mean_by_year_high))
corr_low = rcorr(as.matrix(mean_by_year_low))

corrplot(corr_high$r, is.corr = FALSE, method = "circle")
corrplot(corr_low$r, is.corr = FALSE, method = "circle")

is.nan.data.frame <- function(x)
do.call(cbind, lapply(x, is.nan))
mean_by_year_low[is.nan(mean_by_year_low)] <- 0

```
The output of the function rcorr() is a list containing the following elements :

r : the correlation matrix
n : the matrix of the number of observations used in analyzing each pair of variables
P : the p-values corresponding to the significance levels of correlations.
-------------------------------------------------------------------------------------
Correlation for All Countries 
              Year seatprop ger_tertiary ger_secondary flfpr  gdp  gni
Year          1.00     0.99         0.14          0.20  0.98 0.85 0.98
seatprop      0.99     1.00         0.22          0.29  0.98 0.89 0.98
ger_tertiary  0.14     0.22         1.00          0.82  0.14 0.21 0.13
ger_secondary 0.20     0.29         0.82          1.00  0.14 0.22 0.13
flfpr         0.98     0.98         0.14          0.14  1.00 0.96 0.99
gdp           0.85     0.89         0.21          0.22  0.96 1.00 0.97
gni           0.98     0.98         0.13          0.13  0.99 0.97 1.00

P
              Year   seatprop ger_tertiary ger_secondary flfpr  gdp    gni   
Year                 0.0000   0.5828       0.4258        0.0000 0.0000 0.0000
seatprop      0.0000          0.3698       0.2427        0.0000 0.0000 0.0000
ger_tertiary  0.5828 0.3698                0.0000        0.5826 0.4126 0.6285
ger_secondary 0.4258 0.2427   0.0000                     0.5890 0.3727 0.6269
flfpr         0.0000 0.0000   0.5826       0.5890               0.0000 0.0000
gdp           0.0000 0.0000   0.4126       0.3727        0.0000        0.0000
gni           0.0000 0.0000   0.6285       0.6269        0.0000 0.0000       
-------------------------------------------------------------------------------------
Correlation for High Income 
               Year seatprop ger_tertiary ger_secondary flfpr  gdp   gni
Year           1.00     0.99        -0.03          0.80  0.99 0.85  0.98
seatprop       0.99     1.00         0.00          0.83  0.98 0.85  0.96
ger_tertiary  -0.03     0.00         1.00          0.24  0.00 0.08 -0.05
ger_secondary  0.80     0.83         0.24          1.00  0.72 0.58  0.69
flfpr          0.99     0.98         0.00          0.72  1.00 0.97  0.99
gdp            0.85     0.85         0.08          0.58  0.97 1.00  0.97
gni            0.98     0.96        -0.05          0.69  0.99 0.97  1.00

P
              Year   seatprop ger_tertiary ger_secondary flfpr  gdp    gni   
Year                 0.0000   0.9207       0.0001        0.0000 0.0000 0.0000
seatprop      0.0000          0.9999       0.0000        0.0000 0.0000 0.0000
ger_tertiary  0.9207 0.9999                0.3582        0.9913 0.7400 0.8558
ger_secondary 0.0001 0.0000   0.3582                     0.0018 0.0143 0.0030
flfpr         0.0000 0.0000   0.9913       0.0018               0.0000 0.0000
gdp           0.0000 0.0000   0.7400       0.0143        0.0000        0.0000
gni           0.0000 0.0000   0.8558       0.0030        0.0000 0.0000       
-------------------------------------------------------------------------------------
Correlation for Low Income 
               Year seatprop ger_tertiary ger_secondary flfpr   gdp   gni
Year           1.00     0.97         0.64          0.86 -0.95  0.98  0.99
seatprop       0.97     1.00         0.64          0.80 -0.94  0.97  0.97
ger_tertiary   0.64     0.64         1.00          0.57 -0.64  0.55  0.73
ger_secondary  0.86     0.80         0.57          1.00 -0.75  0.83  0.84
flfpr         -0.95    -0.94        -0.64         -0.75  1.00 -0.98 -0.98
gdp            0.98     0.97         0.55          0.83 -0.98  1.00  1.00
gni            0.99     0.97         0.73          0.84 -0.98  1.00  1.00

P
              Year   seatprop ger_tertiary ger_secondary flfpr  gdp    gni   
Year                 0.0000   0.0073       0.0000        0.0000 0.0000 0.0000
seatprop      0.0000          0.0074       0.0000        0.0000 0.0000 0.0000
ger_tertiary  0.0073 0.0074                0.0222        0.0099 0.0275 0.0022
ger_secondary 0.0000 0.0000   0.0222                     0.0006 0.0000 0.0000
flfpr         0.0000 0.0000   0.0099       0.0006               0.0000 0.0000
gdp           0.0000 0.0000   0.0275       0.0000        0.0000        0.0000
gni           0.0000 0.0000   0.0022       0.0000        0.0000 0.0000       
-------------------------------------------------------------------------------------

#RJAGS MODELING PART


```{r}
library(rjags)

# Normalized datasets
flfpr_rj <- mean_by_year_low$flfpr
seatprop_rj <- mean_by_year_low$seatprop
#ger_tertiary_rj <- mean_by_year_low$ger_tertiary
#ger_secondary_rj <- mean_by_year_low$ger_secondary
gdp_rj <- mean_by_year_low$gdp
gni_rj <- mean_by_year_low$gni
n <- nrow(mean_by_year_low)

# JAGS Little Language
model_string <- "model{

# Likelihood
for(i in 1:n){
mu1[i] <- beta1 + beta2*seatprop_rj[i] + beta3*gni_rj[i]+ beta4*gdp_rj[i]

flfpr_rj[i]   ~ dnorm(mu1[i],inv.var1)
}

# Prior for beta
beta1 ~ dnorm(0,0.01)
beta2 ~ dnorm(0,0.01)
beta3 ~ dnorm(0,0.01)
beta4 ~ dnorm(0,0.01)
#beta5 ~ dnorm(0,0.01)


# Prior for the inverse variance
var1   ~ dunif(0.01, 0.99)
var2   ~ dunif(0.01, 0.99)
var3   ~ dunif(0.01, 0.99)
var4   ~ dunif(0.01, 0.99)
#var5   ~ dunif(0.01, 0.99)

inv.var1   <- 1/var1
inv.var2   <- 1/var2
inv.var3   <- 1/var3
inv.var4   <- 1/var4
#inv.var5   <- 1/var5

sigma1     <- sqrt(var1)
sigma2     <- sqrt(var2)
sigma3     <- sqrt(var3)
sigma4     <- sqrt(var4)
#sigma5     <- sqrt(var5)

}"

model <- jags.model(textConnection(model_string), data = list(flfpr_rj=flfpr_rj, seatprop_rj=seatprop_rj, gdp_rj=gdp_rj, gni_rj=gni_rj,n=n), n.chains = 4)

update(model, 10000, progress.bar="none"); # Burn-in for 10000 samples

samp <- coda.samples(model, 
                     variable.names=c("beta1","beta2","beta3", "beta4", "sigma1", "sigma2", "sigma3","sigma4"), 
                     n.iter=20000, progress.bar="text")

summary(samp)
traceplot(samp)

# sometimes the gelman plot won't fit on a screen
# we have to reduce the margins
par(mar=c(.4,.4,.4,.4))
gelman.plot(samp)
saveRDS(samp, "flfpr_dependency_fork.RDS")

```


```{r}
setwd("~/Documents/GitHub/dav_final")

#graphic1
#LITERACY
YouthLiteracy=read.csv('GenderStatsData/graphic1/youth_literacy_15_24.csv',header = TRUE, sep=',')
length(unique(YouthLiteracy$Country)) #number of countries

#female literacy in 1995 
mean(YouthLiteracy[(YouthLiteracy$Year==min(YouthLiteracy$Year))&(YouthLiteracy$Sex=='Female'),]$Value)
#female literacy in 2016 
mean(YouthLiteracy[(YouthLiteracy$Year==max(YouthLiteracy$Year))&(YouthLiteracy$Sex=='Female'),]$Value)

#MATERNAL MORTALITY per 100,000 births
mortality=read.csv('GenderStatsData/graphic1/maternal_mortality.csv',header = TRUE, sep=',')
#in 1995 
mean(mortality[(mortality$Year==1995)&(mortality$Sex=='Female'),]$Value)
#in 2015 
mean(mortality[(mortality$Year==2015)&(mortality$Sex=='Female'),]$Value)

#EDUCATION 25+ 
uppersecondary=read.csv('GenderStatsData/graphic1/education_uppersecondary_25plus.csv',header = TRUE, sep=',')
#in 1995 
mean(uppersecondary[(uppersecondary$Year==1995)&(uppersecondary$Sex=='Female'),]$Value)
#in 2015 
mean(uppersecondary[(uppersecondary$Year==2015)&(uppersecondary$Sex=='Female'),]$Value)
```

